---
title: "Module 5 -- Skills"
author: "YOUR NAME HERE"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

## Question 1 

The Birdwatchers of America are interested in the populations of bird species common to Western Oregon (data are in Lab4_SkillsData.csv).  For one day, they set up a station and counted the number of American robins, black-capped chickadees and barn swallows that were observed each hour.  

### a)	One of the questions the birdwatchers are interested in is the number of total birds observed in a unit of time.  What discrete distribution would you use for analysis?

```{r}
# Your code here
```

Your final answer here

### b)	What is the expected number of total birds one would see during an hour of bird watching? ‘

```{r}
# Your code here
```

Your final answer here

### c)	What is the probability that in any given hour 6 birds will be observed?  Calculate this probability using both the Poisson probability formula and dpois(). 

```{r}
# Your code here
```

Your final answer here

### d)	What is the probability that you would observe more than 4 birds in any given hour? Calculate this probability using both dpois() and ppois().

```{r}
# Your code here
```

Your final answer here

### e)	Suppose on average you see 2.167 barn swallows and 1.125 American robins per hour. What is the probability that you see 2 or fewer barn swallows or exactly 1 robin in any given hour? Assume that the numbers of swallows and robins are independent.

```{r}
# Your code here
```

Your final answer here

## Question 2

Recall that the birdwatchers observed birds each hour for one day (24 hours). They want to know whether or not a black-capped chickadee was observed each hour.

### a)	What discrete distribution would you use for this analysis?

```{r}
# Your code here
```

Your final answer here

### b)	What proportion of the hours do the observers see a chickadee? 

```{r}
# Your code here
```

Your final answer here

The observers go back into the field a separate day for 8 hours and record whether they see chickadees each hour.  

### c)	Find the mean, variance, and standard deviation for the number of hours in which chickadees will be seen.  

```{r}
# Your code here
```

Your final answer here

### d)	What is the probability within that second, 8-hour observation day that there will be exactly 7 hours in which they observe chickadees? Calculate this probability using both the binomial probability formula and dbinom(). 

```{r}
# Your code here
```

Your final answer here

## Question 3

Assume that the test scores of a college entrance exam fits a normal distribution. Furthermore, the mean test score is 72, and the population standard deviation is 15.2. What is the percentage of students scoring 84 or more in the exam?

```{r}
# Your code here
```

Your final answer here

## Question 4 Use the R pre-packaged dataset “women”,  answering the following questions:

### (a)	What is the z-score that corresponds to a woman with a height of 61 inches? 
```{r}
# Your code here
```

Your final answer here

### (b)	What is the probability of a woman being 61 inches in height or shorter?

```{r}
# Your code here
```

Your final answer here

### (c)	How tall can a woman be and be within the shortest 40% of the population?

```{r}
# Your code here
```

Your final answer here

