---
title: "QSCI 381: Data & Measurement"
author: "Student Name"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

# Overview: Data & Measurement

In this activity, you will:
- Practice working in R and RStudio
- Learn how statisticians think about data
- Distinguish populations from samples
- Identify different types of data
- Think critically about study design and data collection

Use this document to write short answers and run simple R commands.

---

# R & RStudio

## Working in RStudio

Run the code below and answer the questions.

```{r}
getwd()
list.files()
```

**Questions:**
- What is your working directory?
- What files do you see?

---

## Basic R objects

```{r}
x <- 5
y <- 12

x + y
x * y
```

**Question:**  
What is an *object* in R, and why are objects useful?

---

# Introduction to the Field of Statistics

In your own words, answer:

**What is statistics?**

*(1–3 sentences)*

---

# Inference as a Means of Discovery

Statistical inference uses data from a **sample** to learn about a **population**.

**Example inference question (write your own):**
-  

---

# Populations & Parameters vs Samples & Statistics

Fill in the definitions below.

- **Population:**  
- **Parameter:**  
- **Sample:**  
- **Statistic:**  

## Practice

Suppose you measure shell length from 25 oysters at a hatchery.

- Population:
- Sample:
- Parameter:
- Statistic:

---

# Types of Data

Common types of data include:
- **Categorical** (nominal, ordinal)
- **Numerical** (discrete, continuous)

## Classify the following variables

| Variable | Type of Data |
|--------|--------------|
| Species |              |
| Temperature (°C) |     |
| Number of larvae |      |
| Disease status (yes/no) | |
| pH | |

---

# Study Design

A **study design** describes how data are collected to answer a question.

Describe a simple study you could conduct.

- Research question:
- Response variable:
- Explanatory variable(s):
- Units of observation:

---

# Data Collection

List **three potential problems** that can occur during data collection.

1.  
2.  
3.  

---

# Sampling Approaches

Examples of sampling approaches include:
- Random sampling
- Stratified sampling
- Convenience sampling

**Question:**  
Which sampling approach is strongest for making inferences about a population, and why?

---

# Experimental Design

Experiments are used to test **cause-and-effect** relationships.

Key components:
- Treatments
- Replication
- Randomization
- Controls

## Scenario

You test larval survival at three temperatures using multiple tanks per temperature.

Identify:
- Treatments:
- Replicates:
- Response variable:
- Why this is an experiment:

---

# Data & Measurement: Summary & Application

## Create a small dataset

```{r}
data <- data.frame(
  temperature = c(10, 10, 14, 14, 18, 18),
  survival = c(0.85, 0.88, 0.75, 0.78, 0.60, 0.62)
)

data
```

## Summarize the data

```{r}
mean(data$survival)
sd(data$survival)
```

**Questions:**
- What does the mean represent?
- Is this a statistic or a parameter?
- What inference question could you ask with this data?

---

# Reflection

In 2–3 sentences:
- What was the most important concept from today?
- What is still confusing?
