---
title: "QSCI 381: Data & Measurement -- KEY"
author: "Student Name"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

# Overview: Data & Measurement

In this activity, you will:
- Practice working in R and RStudio
- Learn how statisticians think about data
- Distinguish populations from samples
- Identify different types of data
- Think critically about study design and data collection

Use this document to write short answers and run simple R commands.

---

# R & RStudio

## Working in RStudio

Run the code below and answer the questions.

```{r}
getwd()
list.files()
```

**Questions:**
- What is your working directory?

The **working directory** is the folder (or "directory") on your computer where R is currently reading and saving files. It is the default location, and other file locations are described in reference to the working directory.

- What files do you see?

You should see the `01-Skills.Rmd` file.

---

## Basic R objects

```{r}
x <- 5
y <- 12

x + y
x * y
```

**Question:**  
What is an *object* in R, and why are objects useful?

An R **object** is a named container that stores values or data (for example, a vector or a data frame). Objects are useful because they allow data to be saved, reused, manipulated, and referenced throughout an analysis.

---

# Introduction to the Field of Statistics

In your own words, answer:

**What is statistics?**

Statistics is the science of collecting, analyzing, and interpreting data in order to learn about the world and make informed decisions under uncertainty.

---

# Inference as a Means of Discovery

Statistical inference uses data from a **sample** to learn about a **population**.

**Example inference question (write your own):**

Ex. "Is the average shell length of oysters at this hatchery different from the historical average?"

---

# Populations & Parameters vs Samples & Statistics

Fill in the definitions below.

- **Population**:
The entire group of individuals or observations of interest.

- **Parameter**:
A numerical value that describes a characteristic of a population.

- **Sample**:
A subset of the population that is actually measured or observed.

- **Statistic**:
A numerical value calculated from a sample that summarizes the data. 

## Practice

Suppose you measure shell length from 25 oysters at a hatchery.

- Population:

All oysters in the hatchery

- Sample:

The 25 oysters measured

- Parameter:

True mean shell length of all oysters at the hatchery (the population)

- Statistic:

The mean shell length calculated from the 25 measured oysters (the sample)

---

# Types of Data

Common types of data include:
- **Categorical** (nominal, ordinal)
- **Numerical** (discrete, continuous)

## Classify the following variables

| Variable | Type of Data |
|--------|--------------|
| Species | Categorical (nominal) |
| Temperature (°C) | Numerical (continuous) |
| Number of larvae | Numerical (discrete) |
| Disease status (yes/no) | Categorical (nominal) |
| pH | Numerical (continuous) |

---

# Study Design

A **study design** describes how data are collected to answer a question.

Describe a simple study you could conduct.

- Research question:

"Does temperature affect larval survival in oysters?"

- Response variable:

Larval survival rate.

- Explanatory variable(s):

Temperature.

- Units of observation:

Individual larvae (or tanks, depending on measurement).

---

# Data Collection

List **three potential problems** that can occur during data collection.

1. Measurement error or inconsistent measurement methods
2. Biased or non-random sampling
3. Missing, lost, or incorrectly recorded data

---

# Sampling Approaches

Examples of sampling approaches include:
- Random sampling
- Stratified sampling
- Convenience sampling

**Question:**  
Which sampling approach is strongest for making inferences about a population, and why?

Generally, **random sampling** is the strongest because it reduces bias and ensures that the sample is representative of the population, allowing valid statistical inference.

---

# Experimental Design

Experiments are used to test **cause-and-effect** relationships.

Key components:
- Treatments
- Replication
- Randomization
- Controls

## Scenario

You test larval survival at three temperatures using multiple tanks per temperature.

Identify:
- Treatments:

**Temperature** (three levels)

- Replicates:

The multiple tanks at each temperature

- Response variable:

Larval survival

- Why this is an experiment:

The treatment, temperature, is being **actively manipulated** by the researcher, allowing cause-and-effect conclusions to be tested.

---

# Data & Measurement: Summary & Application

## Create a small dataset

```{r}
data <- data.frame(
  temperature = c(10, 10, 14, 14, 18, 18),
  survival = c(0.85, 0.88, 0.75, 0.78, 0.60, 0.62)
)

data
```

## Summarize the data

```{r}
mean(data$survival)
sd(data$survival)
```

**Questions:**
- What does the mean represent?

The mean represents the average value of the measured variable across all observations in the datset. It summarizes central tendency, but is heavily influenced by extreme values.

- Is this a statistic or a parameter?

We calculated the mean from a *sample*, rather than from an entire population. A parameter would describe a true population value.

- What inference question could you ask with this data?

Ex. "Is the mean larval survival in this population greater than 0.75?"

---

# Reflection

In 2–3 sentences:
- What was the most important concept from today?

Any reasonable, thoughtful answer

- What is still confusing?

Any reasonable, thoughtful answer
