---
title: "Module 3 Homework: Descriptive Statistics & Probability"
author: "YOUR NAME HERE"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

This assignment has two parts. In the first section you'll work with a
medical data set where patients were randomly assigned to two anesthetic
treatment groups, and the primary outcome was the measured time from
administering the anesthetic to when the anesthetics took full effect.
For this scenario you'll calculate descriptive statistics and create a
visualization.

In the second part, former QSCI 381 students data about their dominant
eye, and where they sat in the lecture auditorium. With this data you'll
be creating contingency tables and calculating probabilities for certain
events.

OBJECTIVES

-   Calculate descriptive statistics for a given dataset.

<!-- -->

-   Calculate and interpret z-scores.

<!-- -->

-   Create visualizations including a box-whisker plot for a given
    dataset

-   Construct a contingency table

-   Calculate empirical probabilities

-   Calculate expected probabilities of occurrence for given
    combinations of events.

# PART A

In this question we will be using a medical dataset examining the
effectiveness of two different methods of anesthetic delivery.

```{r}
# Change file path if necessary
med.data <- read.csv("anesthetic.csv")
```

This data set contains 103 patients who were scheduled to undergo an
upper extremity procedure. For these procedures two different types of
anesthetic are used (ropivacaine, mepivacaine), and the researchers
wished to identify whether it was more advantageous to administer both
at the same time (mixture), or to administer one anesthetic first,
followed by the second after a period of 5 minutes (sequential
treatment). Patients were randomly assigned among the two treatment
groups, and the primary outcome was the measured time from administering
the anesthetic to when the anesthetics took full effect (4-nerve sensory
block onset).

The data consists of the following variables for each patient

-   subject: patient identifier number

-   group: treatment group they were assigned to -- Mixture (group = 1)
    or Sequential (group = 2)

-   gender: patient gender -- Female (gender = 0), Male (gender = 1)

-   bmi: body mass index of the patient (numeric values)

-   age: patient age in years

-   onset_sensory: time until anesthetic took effect (minutes)

## 1)

For each of the two treatment groups calculate the

-   Sample size

-   Mean

-   Median

-   Minimum

-   Maximum

-   Standard Deviation

of the time taken until the anesthetic took effect (i.e. of
onset_sensory) and place your results into a table with one column for
each treatment group and a row for each descriptive statistic, making
sure your rows are labeled. Round all of your answers to one decimal
place (12 pts).

You can create this table using R code, or by switching to the "Visual"
tab of your R Markdown and inserting a table.

```{r}
# Your code here
```

Your final answer(s) here

## 2)

For each treatment group calculate the coefficient of variation. Show
your answers below, rounding numeric values to two decimal places (4
pts)

```{r}
# Your code here
```

Your final answer(s) here

## 3)

For each treatment group calculate the:

-   lower (Q1) quartile

-   upper (Q3) quartile,

-   the interquartile range (IQR)

-   upper decile (i.e. 90% percentile)

-   lower decile (i.e. 10% percentile)

```{r}
# Your code here
```

Your final answer(s) here

## 4)

Which treatment group had the largest IQR? Compare and contrast this
with the coefficient of variation finding from Question 2? (12 pts).

```{r}
# Your code here
```

Your final answer(s) here

## 5)

Using the boxplot() or plot() function, plot anesthetic response times
(onset_sensory) as a box-whisker plot with onset_sensory on the y-axis,
and treatment group on the x-axis. Label the axes accordingly (including
units for onset_sensory), and include your name as the title of the
plot. (5 pts)

```{r}
# Your code here
```

## 6)

Based on the mean and median response times, and via examination of the
boxplots (created in Question 5), is the distribution of response times
symmetric, left-skewed or right-skewed, giving reasons for your answer
(3 pts).

Your final answer(s) here

## 7)

Calculate the z-scores for a response time of 20 minutes for each
treatment group based on each group's mean and standard deviation. Round
your answers to 2 decimal places (4 pts)

```{r}
# Your code here
```

Your final answer(s) here

## 8)

Using the mean and standard deviation calculated for each group,
calculate the proportion of observations for each group that are within
± 1 SD of the mean, and also the proportion that are within ± 2 SD of
the mean. Round your answers to 3 decimal places

```{r}
# Your code here
```

Your final answer(s) here

## 9)

Compare these proportions to the expected values according to the
Empirical Rule, noting any differences between the calculated values,
and those expected from the Empirical Rule, and why this difference may
have occurred (12 pts)

Your final answer(s) here

# PART B

In this question we will work with contingency tables and calculate
probabilities, using the eye dominance dataset we collected in class.

```{r}
# Change file path if necessary
eyes <- read.csv("QSCI381Eyes.csv")
```

This dataset consists of survey results regarding eye dominance, and
seat location within the lecture theatre. The data consists of the
following variables:

-   ID: Participant ID number

-   eye: whether the individual was right or left eye dominant

-   side: the side of the room the participant was sat in

-   row: the row number the individual was sat in

-   morning: whether the participant reported as being a morning person
    or not

## 10)

Construct a contingency table of the frequency of eye dominance (2 rows;
right or left) by which side of the room they sat in (3 columns; Left,
Middle, Right). Make sure you label rows and columns. Include column and
row totals. (8 pts)

You can create this table using R code, or by switching to the "Visual"
tab of your R Markdown and inserting a table.

```{r}
# Your code here
```

Your final answer(s) here

## 11)

Calculate the empirical probabilities for the following scenarios

-   The probability an individual is left eye dominant

-   The probability an individual is right-eye dominant AND sat in the
    middle of the room

-   The probability an individual is left eye dominant GIVEN they are
    sat in the left side of the room

-   The probability an individual is right-eye dominant OR is sat in the
    right side of the room

In each case round your answer to 3 decimal places (8 pts)

```{r}
# Your code here
```

Your final answer(s) here

## 12)

Using the probability an individual is right-eye dominant P(right-eye),
and the probability an individual sat in the left of the room
P(left-room), calculate the expected probability of an individual being
right-eye dominant AND that sat in the left side of the room P(right-eye
AND left-room) assuming that these events are Independent using the
Multiplication Rule. Round your answer to 3 decimal places (4 pts).

```{r}
# Your code here
```

Your final answer(s) here

## 13)

Now calculate the empirical observed probability of an individual being
right eye dominant AND that sat in the left side of the room. Round your
answers to 3 decimal places (2 pts)

```{r}
# Your code here
```

Your final answer(s) here

## 14)

Comparing your empirical probability (calculated in 13) to the expected
probability (calculated in 12c), do you think the side of the room sat
in and eye dominance are independent? (1 pt)

Your final answer(s) here
