---
title: "Module 02 – KEY"
date: "2026-01-12"
output: html_document
---

## Overview

We will calculate descriptive statistics of GPS data describing the location of an invasive bird species on the island of Nusa Penida, Indonesia. Specifically, we will examine the *Horizontal Dilution of Precision* (HDOP) for each location recording. HDOP is an indicator of accuracy for each recording (higher values indicate lower accuracy).

## Setup

```{r setup, include=TRUE}

# read in data (edit the path as needed)
gps <- read.csv('GPS.csv', header = TRUE)

# view data (optional)
 View(gps)
```

## (1a) Mean HDOP

Calculate the mean for the column of data labeled `HDOP`.

```{r}
mean(gps$HDOP)
```

Expected output:

```
2.611111
```

## (1b) Population or sample mean?

This is the **mean of the sample**.

## (2) Minimum, maximum, and range of HDOP

Calculate the minimum, maximum, and range for the column of data labeled `HDOP`.

```{r}
min(gps$HDOP)
max(gps$HDOP)

r_h <- range(gps$HDOP)
r_h[2] - r_h[1]
```

Expected output:

- min: `1`
- max: `12.7`
- range: `11.7`

## (3) Median HDOP

Calculate the median for the column of data labeled `HDOP`.

```{r}
median(gps$HDOP)
```

Expected output:

```
1.7
```

## (4) Standard deviation of HDOP

Calculate the standard deviation for the column of data labeled `HDOP`.

```{r}
v_h <- var(gps$HDOP) # The denominator n - 1 is used by this function
sqrt(v_h)

sd(gps$HDOP) # Like var, this uses denominator n - 1.
```

Expected output:

```
2.764101
```

## (5) Histogram of HDOP

Create a histogram for the column of data labeled `HDOP`. Include a title and x and y axis labels.

```{r}
hist(gps$HDOP,
     main = 'frequence of the HDOP',
     xlab = 'Horizontal Dilution of Precision',
     ylab = 'frequencies')
```
