---
title: "Module 02 Skills"
author: 
date: "`r Sys.Date()`"
output: pdf_document
---

We will be calculating descriptive statistics of some GPS data describing the location of an invasive bird species on the island of Nusa Penida, Indonesia. Specifically, we will look at the Horizontal Dilution of Precision (or HDOP) of each location recording. The HDOP is an indicator of the accuracy for each recording (higher values indicate lower accuracy). We’d like to explore some statistics to get an idea of the quality of the GPS data.

Download “GPS.csv” from the course Canvas site and read it into R using the read.csv() command (be sure to set your working directory to the file’s location! The folder that your csv file is in)

```{r}
# Your code here
```

Your final answer here

## (1a) 

Calculate the mean for the column of data labeled HDOP.

```{r}
# Your code here
```

Your final answer here

## (1b) 

Is this the mean of the population or the mean of the sample? 

Your final answer here

## (2) 

Calculate the minimum, maximum, and range for the column of data labeled HDOP.


```{r}
# Your code here
```

Your final answer here

## (3) 

Calculate the median for the column of data labeled HDOP.

```{r}
# Your code here
```

Your final answer here

## (4) 

Calculate the standard deviation for the column of data labeled HDOP.

```{r}
# Your code here
```

Your final answer here

## (5) 

Create a histogram for the column of data labeled HDOP. Include a title and x and y axis labels.

```{r}
# Your code here
```

Your final answer here

